---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("data", "CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
levels(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.
```{r}
crabs <- crabs |> filter(Sex == "F")
```


2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

#We will use a regression test because we are looking for a relationship between a predictor and a response variable and we only have one predictor variable.

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

Plot the data
```{r}
ggplot(crabs, aes(x=Weight, y = Height)) +
  geom_point()
```
There appears to be a strong correlation between weight and height

Make a linear model of the data
```{r}
crabs_model <- lm(Height ~ Weight, data = crabs)
```

Use autoplot to look at assumptions
```{r}
autoplot(crabs_model, smooth.colour = NA)
```


Interpretation: crabs who weigh more will be significantly taller. (F = 613.6, DF = 1223, P = < 2.2x10^-16)

create a summary table
```{r}
summary(crabs_model)
```
Crabs who weigh more will be significantly taller ()


Create a better table
```{r}
ggplot(crabs_model, aes(x = Weight,
y = Height)) +
geom_point() +
geom_smooth(method = 'lm') +
theme_bw()
```

## 
