---
title: "Homework-26"
format: html
editor: visual
---

Set up:
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(multcomp)
```

Import data set:
```{r}
pea <- read.csv(here("data", "s_r_pea_sections.csv"))
```

Convert Treatment from a character to a factor variable:
```{r}
pea$Treatment <- as.factor(pea$Treatment) 
```


Create a plot:
```{r}
ggplot(pea, aes(x=Treatment, y = pea_section_length_ou)) +
  geom_boxplot() +
  coord_flip() + # makes treatment types readable
  geom_point() + # adds points to the boxplot
  geom_jitter(width = .25) # makes all points visible
```

initial interpretation:
Control treatment has the highest mean length and the mix of fructose and glucose has the lowest mean length. 2% glucose, 2% fructose and the mix all appear to have have similar means


Make linear model:
```{r}
pea_model <- lm(pea_section_length_ou ~ Treatment, data = pea)
```


Test assumptions:
```{r}
autoplot(pea_model, smooth.colour = NA)
```
Figures suggest that model is okay


Conduct a tukey HSD test and interpret results:
```{r}
pea_tukey <- glht(pea_model, linfct = mcp(Treatment = "Tukey"))
summary(pea_tukey) # show results of tukey test
```
2% sucrose and the mix have a significant difference in mean pea length. Control and the mix have a significant difference in mean pea length. 2% sucrose and 2% fructose have a significant difference in mean pea length. Control and 2% fructose have a significant difference in mean pea length. 2% sucrose and 2% glucose have a significant difference in mean pea length. Control and 2% glucose have a significant difference in mean pea length. Control and 2% sucrose have a significant difference in mean pea length.


