---
title: "Homework-21"
format: html
editor: visual
---

Trying a chi-square contingency analysis:

```{r}
rm(list = ls())
# clear the environment

library(tidyverse)
library(here)
# import packages

lady <- read.csv(here("data", "ladybirds.csv"))
# import the data

glimpse(lady)
# look at the data
```

convert character variables to factors

```{r}
lady$Habitat <- as.factor(lady$Habitat)
lady$Site <- as.factor(lady$Site)
lady$colour <- as.factor(lady$colour)
```

Calculate total number of ladybirds based on color and habitat

```{r}
totals <- lady |>
  group_by(Habitat, colour) |>
  summarise(total_number = sum(number))
# group the totals by their habitat and color
```

create a bar chart using the totals

```{r}
ggplot(totals, aes(x = Habitat, y = total_number, fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge')
# creates a bar graph that has a column for both colors within both industry. columns are color-coded based on color of ladybirds
```

Recreate the above graph with better colors

```{r}
ggplot(totals, aes(x = Habitat, y = total_number,
fill = colour)) +
geom_bar(stat = 'identity', position = 'dodge') +
scale_fill_manual(values = c(black = "black", red = "red"))
# change black color to black and red color to red
```

Turn the columns into a matrix with color and habitat

```{r}
lady_mat <- xtabs(number ~ Habitat + colour,
data = lady) 
# `xtabs()` means crosstab totals by habitat and color
```

run the chi-squared test

```{r}
chisq.test(lady_mat)
# use `chisq.test()` on the lady_mat matrix

lady_chi <- chisq.test(lady_mat)
names(lady_chi)
lady_chi$expected
# returns the expected values
```

T-TEST:

```{r}
ozone <- read.csv(here("data", "ozone.csv"))
# import the data frame

glimpse(ozone)
# look at structure of the DF

ozone$Garden.location <- as.factor(ozone$Garden.location)
ozone$Garden.ID <- as.factor(ozone$Garden.ID)
# turn character variables into factors
```

Plot the data as a histogram

```{r}
ggplot(ozone, aes(x = Ozone)) +
geom_histogram(binwidth = 10) +
facet_wrap(~ Garden.location, ncol = 1) +
theme_bw()
# stacking the graphs can help us see the difference between the means
```

Complete a t-test

```{r}
t.test(Ozone ~ Garden.location, data = ozone)
# shows the tscore, degrees of freedom, p-value < .05, mean in East is 77.34 and mean in west is 61.26
```

complete a variance test using var.test()

```{r}
var.test(Ozone ~ Garden.location, data = ozone)
# not necessary when running the welch version
```
