---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

```{r}
df_No_NA <- df[complete.cases(df), ]
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

There are 2 DF because 2 of the treatments can vary.

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this:

There are 46 DF because the complete rows - the treatments is how you find the DF for residuals.

```{r}
sum(!is.na(df$Growth_rate_mm))
```

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

49 - 3 = 46 = degrees of freedom for the residuals.

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

No, the p-value is higher than .05 which means that there is no significant difference between treatment 1 and control.

Does adding 2 doses of fertilizer change the growth rate significantly from the Control?

Yes, the p-value is smaller than .05 which means there is a significant difference between treatment 2 and control.

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(x = Treatment, y = Growth_rate_mm)) +
  geom_boxplot() +
  geom_point() +
  theme_bw()
```

What do you think?

I believe that there will be a significant difference.

#### Create a model:

```{r}
grow.mod.no.NA <- lm(Growth_rate_mm ~ Treatment, data = df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA?

They look pretty similar to our first anova.

#### Interpret results:

```{r}
anova(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

There are 2 DF because there are 2 treatments that can vary.

Why are there 33 DF for the Residuals?

```{r}
sum(!is.na(df_No_NA$Growth_rate_mm))
```

There are 33 DF because there are only 36 points used and you find the residual DF by subtracting the 3 treatments.

How would you write a results statement?

We do not reject our null hypothesis (p-value \> .05). There is no significant difference between control, treatment 1, or treatment 2.

```{r}
summary(grow.mod.no.NA)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

No, it does not significantly change growth rate.

Does adding 2 doses of fertilizer change the growth rate signficantly from the Control?

yes, it does significantly change growth rate.

We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(df_All_NA, aes(x = Dried_mass_g, y = Growth_rate_mm)) +
  geom_point()
```

#### Create model

```{r}
reg_all_model <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_All_NA)
```

#### Check assumptions

```{r}
autoplot(reg_all_model)
```

Don't worry if assumptions don't look perfect - focus on normality of the data.

#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(reg_all_model)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals?

There is only 1 degree of freedom because there is only 1 factor that can vary. There are 34 residual DF because there are only 36 points used and you subtract the 2 treatments from 36 to get 34. 
There are only 34 variables because only 34 variables are complete in growth rate and dried mass columns.

```{r}
summary(reg_all_model)
```

There is a relationship between dried mass and growth rate (p-value \< .05)

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(df_No_NA, aes(x=Dried_mass_g, y = Growth_rate_mm)) +
  geom_point()
```

#### Create model

```{r}
reg_no_model <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_No_NA)
```

#### Check assumptions

```{r}
autoplot(reg_no_model)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis?

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(reg_no_model)
```

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why?

1 df because there are only 2 treatments and 1 can vary

2.  How many degrees of freedom are there for Residuals? Why?

34 because 36 points are used and you subtract the 2 treatments from the 36.

3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis?

The DF are the same in both models.
