---
title: "Homework-25"
format: html
editor: visual
---

Set up qmd
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

import and look at the data
```{r}
daphnia <- read.csv(here("data", "Daphniagrowth.csv"))

glimpse(daphnia)
```

turn parasite into a factor
```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```


create an initial graph of the data
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
geom_boxplot() +
  theme_bw() + 
  coord_flip() # new function that flips the x and y axis so that we can see the full parasite name
```

We can see that the growth rates vary between the treatments, control has the highest growth rate, and there is likely an ordering in the treatments.

construct a linear model:
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

evaluate the data using `autoplot()`
```{r}
autoplot(model_grow, smooth.colour=NA)
```
All of figures suggest that everything is okay.


implement the anova test
```{r}
anova(model_grow)
```
Our F value is large because there is such a big difference between the largest and smallest treatments. This also results in a smaller p-value. This means we can reject the null hypothesis.


Check the summary table
```{r}
summary(model_grow)
```

Another way to show the means for each treatment level:
```{r}
sumDat <- daphnia |> 
  group_by(parasite) |> 
  summarise(meanGR = mean(growth.rate))
sumDat
```

Create a new graph
```{r}
ggplot(daphnia, aes(x = growth.rate, y = parasite)) +
  geom_point(alpha = .3, aes(color=parasite), size =4) +
  geom_point(data = sumDat, aes(meanGR, parasite, color=parasite), shape = "diamond", size = 8)
```

