---
title: "HW-28"
format: html
editor: visual
---

In this file, I will be going over different graphing techniques

Import the data and packages
```{r}
library(tidyverse)
library(here)
library(gridExtra)

compensation <- read.csv(here("data", "compensation.csv" ))
```

Create a basic scatterplot and box and whisker plot
```{r}
eg_scatter <- ggplot(data = compensation, aes(x=Grazing, y= Fruit)) +
  geom_point() # assigns scatterplot to eg_scatter


eg_box <- ggplot(data = compensation, aes(x=Grazing,y=Fruit)) +
  geom_boxplot() # assigns boxplot to eg_box
```

add a theme to the plot
```{r}
eg_scatter+theme_bw()
```

Place both plots on one page
```{r}
grid.arrange(eg_scatter, eg_box, nrow=1)
```

Use `xlab()` and `ylab()` to adjust axis
titles
```{r}
eg_scatter + xlab("Root Biomass") + ylab("Fruit Production")
```

Use `ggtitle()` to add a title to the plot but use sparingly
```{r}
eg_scatter + ggtitle("My title")
```

We can combine all of these using `labs()`
```{r}
eg_scatter + labs(title="title", x="Root Biomass", y="Fruit Production")
```

We can change the range of x and y
```{r}
eg_scatter + xlim(0,20) + ylim(0,140)
```

`annotate()` allows you to place text inside the plot
```{r}
eg_scatter +
  annotate("text", x = c(6,8), y=c(105,25),
           label=c("text up here","text down here")) # specify the location of the text using x and y coords then write message
```

Adjust the scales
```{r}
eg_scatter + scale_x_continuous(limits = c(4,11), breaks= 4:11)
```

Change colors
```{r}
ggplot(data = compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
geom_point() +
scale_colour_manual(values = c(Grazed = "brown", Ungrazed =
"green"))
```

Adjust background and grids
```{r}
eg_scatter +
theme(
panel.background = element_rect(fill = NA, colour = "black"),
panel.grid.minor = element_blank(),
panel.grid.major = element_line(colour = "lightblue")
)
```

adjust color and size of x axis and angle of x axis tick label
```{r}
eg_box +
theme(
axis.title.x = element_text(colour = "cornflowerblue",
size =rel(2)),
axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))
```

Change names of levels on our plot
```{r}
eg_box + scale_x_discrete(limits = c("Ungrazed", "Grazed"),
labels = c("Control", "Grazed"))
```

look at boxes using clean key
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
geom_point() +
theme(legend.key = element_rect(fill = NA))
```

