---
title: "Homework 23"
format: html
editor: visual
---

We are looking into linear models which include regression, ANOVA, ANCOVA

Set up:
```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Import the data set:
```{r}
plant_gr <- read.csv(here("data", "plant.growth.rate.csv"))
```

Examine the data:
```{r}
glimpse(plant_gr)
```

Create a scatterplot of the data:
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() + ylab("Plant Growth Rate (mm/week)") +
theme_bw()
```
The slope is positive which means the more moisture thats in the soil, the higher the plant growth rate is


Use `lm()` to fit the model:
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, data = plant_gr) # using plant growth rate as a function of soil moisture content
```


Use autoplot function to plot the linear model:
```{r}
library(ggfortify)
autoplot(model_pgr, smooth.color = NA)
```
1) the top left graph tells us if there is a line that fits the data

2) the top right tells us that the residuals as dots and the expectation when conditions are normal

3) the bottom left tells shows us what the assumption of equal variance looks like. See if theres a pattern.

4) the bottom right tells us that shows influential points and outliers

The `smooth.color = NA` removes a line that is fitted based on locally weighted regression because it typically does not look at the data points


`anova()` does not perform the ANOVA test that compares means

`summary()` produces a table of estimates based on the model

```{r}
anova(model_pgr)
```
shows us a large F value and a tiny p value. If we run it again and get similar values we can be pretty confident that the pattern is correct

```{r}
summary(model_pgr)
```
the slope is associated with the explanatory variable

create a graph using our model
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content,
y = plant.growth.rate)) +
geom_point() +
geom_smooth(method = 'lm') +
ylab("Plant Growth Rate (mm/week)") +
theme_bw()
```

`geom_smooth()` adds a fitted line to the graph with teh error appearing as a gray background


